Co-citations in context

### 2019-06-06 cocit_r.tex
minor revisions

\author[Author Names]
{James Bradley\affil{1},
Sitaram Devarakonda\affil{2}, Avon Davey\affil{2}, Dmitriy Korobskiy\affil{2}, Siyu Liu\affil{2}, Tandy Warnow\affil{3}
\and George Chacko\affil{2}}

@@

\keywords{(bibliometrics, co-citation analysis, random graphs )}

@@

Key to Uzzi {\em et al.}, however, is their random graph model and its underlying assumptions. A careful examination of their Monte Carlo simulations reveals that random selection of the references used to generate a null model and expected values is equiprobable with respect to disciplinary origin and citation count.  For example, the model permits substitution of a reference in quantum physics with equal probability by a reference in the related quantum chemistry field or in some entirely different field, such as classical literature, evolutionary biology, or anthropology. Such substitutions poorly model documented citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979}. In addition, under this random model, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once, which is inconsistent with citation behavior and inconsiderate of the power law or lognormal citation distributions described in the literature~\citep{stringer_statistical_2010,perline_strong_2005}.  
Lastly,  as we will show, while the citation switching algorithm used in Uzzi  {\em et al.} preserves the number of publications and the number of references in each publication, it is not designed to preserve disciplinary proportions of cited references within subsets of the Web of Science. Accordingly, the expected value calculations generated by the \cite{uzzi_atypical_2013} simulations and used in characterizing journal pairs in terms of conventionality and novelty  can be reasonably questioned on grounds of model mis-specification.

@@

@@

\begin{figure}
\centering
\includegraphics[width=0.5\hsize]{fold_effect.pdf}
\caption{\label{fold_effect} Intra-network citation shuffling preserves disciplinary composition. Publications of type Article belonging to the WoS research areas (i) Arts and Humanities (ii) Life Sciences and Biomedicine were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). The disciplinary composition of cited reference before and after shuffling was measured for each of 151 sub-groups (from the extended subject classification in WoS) in these two research areas, expressed as fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. A fold difference of 1 indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All four boxplots are generated from 151 observations. Min \& Max values for fold differences were as follows. (i) Arts \& Humanities. bg\_local (1,2), bg\_allWoS (1, 488) (ii) Life Sciences \& Biomedicine. bg\_local(1,1), bg\_allWos (1, 352) Note y-axis: $log_2$ scale.} 
\end{figure}

>>>>

\author[Author Names]
{James Bradley\affil{1},
Sitaram Devarakonda\affil{2}, Avon Davey\affil{2}, Dmitriy Korobskiy\affil{2}, Siyu Liu\affil{2}, Djamil Lakhdar-Hamina\affil{2},Tandy Warnow\affil{3}
\and George Chacko\affil{2}}

@@

\keywords{(co-citation analysis, bibliometrics, random graphs )}

@@

Key to Uzzi {\em et al.}, however, is their random graph model and its underlying assumptions. A careful examination of their Monte Carlo simulations reveals that random selection of the references used to generate a null model and expected values is equiprobable with respect to disciplinary origin and citation count.  For example, the model permits substitution of a reference in quantum physics with equal probability by a reference in the related quantum chemistry field or in some entirely different field, such as classical literature, evolutionary biology, or anthropology. Such substitutions poorly model documented citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979}. In addition, under this random model, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once, which is inconsistent with citation behavior and inconsiderate of the power law or lognormal citation distributions described in the literature~\citep{stringer_statistical_2010,perline_strong_2005}.  
Lastly,  as we will show, while the citation switching algorithm used in Uzzi  {\em et al.} preserves the number of publications and the number of references in each publication, it is not designed to preserve disciplinary proportions of cited references within subsets of the Web of Science. Accordingly, the expected value calculations generated by the \cite{uzzi_atypical_2013} simulations and used in characterizing journal pairs in terms of conventionality and novelty  can be reasonably questioned on grounds of model mis-specification when subsets of the WoS are studied.

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=0.8\linewidth]{background-effect}     
\caption{\ Intra-network citation shuffling preserves the disciplinary composition of references. Publications of type Article belonging to the three disciplinary networks (i) Applied Physics (ii) Immunology (iii) Metabolism were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). Citation shuffling was performed using either our algorithm (repcs) or that of Uzzi (umsj). The disciplinary composition of cited references before and after shuffling was measured as frequencies for each of 153 sub-disciplines (from the extended subject classification in WoS) and expressed as fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. Null values were set to 1. A fold difference of 1 indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All four boxplots are generated from 151 observations. Min \& Max values for fold differences were as follows. (i) Arts \& Humanities. bg\_local (1,2), bg\_allWoS (1, 488) (ii) Life Sciences \& Biomedicine. bg\_local(1,1), bg\_allWos (1, 352) Note y-axis: $log_2$ scale.} 
\label{be}
\end{figure}

@@
